Fast multipoint evaluation for exponential size CNFs and SYM$\circ$SYM circuits

### 2022-06-10 sym_eval/main.tex
Symsym cont

\newcommand{\NN}{\mathbb{N}}%
\newcommand{\ZZ}{\mathbb{Z}}%
\newcommand{\Sbb}{\mathbb{S}}%
\newcommand{\Cc}{\mathcal{C}}%
\newcommand{\Gg}{\mathcal{G}}%
\newcommand{\Ff}{\mathcal{F}}%
\newcommand{\Ss}{\mathcal{S}}%
\newcommand{\eps}{\varepsilon}%
\newcommand{\epsm}{\eps^{-1}}%
\newcommand{\bit}{\{0,1\}}%
\DeclareMathOperator*{\poly}{poly}

@@

\begin{theorem}
	For all $n, m$, \ssym{} circuits of size $m$ over $n$ variables can be evaluated
	on every input in time $O^*(2^n + m)$.
\end{theorem}

@@

\begin{theorem} (??)
	With our rules, we cannot do better than $4 \cos(\pi/7)$.
\end{theorem}

@@

A symmetric function is a boolean function whose output only depends on the number of 
its input that equal to 1.
More formally:
\begin{definition}[Symmetric functions]
	A boolean function $f: \bit^n \mapsto \bit$ is \textit{symmetric} if there exists a function
	$g: \{0,\ldots, n\} \mapsto \bit$ such that for every $x\bit^n$:
	\[f(x) = g(\sum_{i=1}^n x_i).\]
\end{definition}

@@

wlog, \ssym{} circuit only contains so-called ``Exact-Majority'' gates.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\bibliography{biblio.bib}

>>>>

\newcommand{\FF}{\mathbb{F}}%
\newcommand{\NN}{\mathbb{N}}%
\newcommand{\ZZ}{\mathbb{Z}}%
\newcommand{\Sbb}{\mathbb{S}}%
\newcommand{\Cc}{\mathcal{C}}%
\newcommand{\Gg}{\mathcal{G}}%
\newcommand{\Ff}{\mathcal{F}}%
\newcommand{\Ss}{\mathcal{S}}%
\newcommand{\eps}{\varepsilon}%
\newcommand{\epsm}{\eps^{-1}}%
\newcommand{\bit}{\{0,1\}}%
\DeclareMathOperator*{\poly}{poly}

@@

\begin{theorem}
	For all $n, m$, \ssym{} circuits of size $m$ over $n$ variables, with possibly negated inputs, can be evaluated
	on every input in time $O^*(2^n + m)$.
\end{theorem}

@@

@@

A symmetric function is a boolean function whose output only depends on the number of 
its input that equal to 1.
More formally:
\begin{definition}[Symmetric functions]
	A boolean function $f: \bit^n \rightarrow \bit$ is \textit{symmetric} if there exists a function
	$g: \{0,\ldots, n\} \rightarrow \bit$ such that for every $x\in\bit^n$:
	\[f(x) = g\left(\sum_{i=1}^n x_i\right).\]
\end{definition}
\todo[inline]{maybe it's easier to write everything over -1,1 ?}

@@

For a symmetric function $f$ and its associated function $g$, we define $T(f) = g^{-1}(1)$,
that is, the set of integers $t$ such that $f$ is one when exactly $t$ inputs of $f$ are equal to 1.

Additionnally, for a gate $f$, we define $S^+(f)$ (resp. $S^-(f)$)
to be the set of variables that are positive (resp. negative) inputs to $f$.

\todo[inline]{What if a variables goes TWICE into a gate ? Doesn't matter for CNFs, but otherwise this needs to be a multiset, and we would have $X_i^{2a}$}

wlog, \ssym{} circuit only contains so-called ``Exact-Majority'' gates.
I don't even need this though.

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Encoding \ssym{} as polys}
\[\sum_{a = 1}^q \sum_{t = -n}^n g^{-at} \sum_{f : t\in T(f)} (\prod_{i \in S^+(f)} X_i^a) (\prod_{i \in S^-(f)} X_i^{-a})\]

1. show that $P(x) = $ number of bottom layer gates equal to 1.

Evaluate over $\{g, g^{-1}\}^n$ in $\FF_p$, where $q\mid p-1$.
How to find $p,q$ so that we don't blow up the running time.

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast multipoint evaluation of these polys.}
Since $q, t = O(n)$, we can iterate over outer loops:
\[ P = \sum_{a = 1}^q \sum_{t = -n}^n g^{-at} P_{a,t}(X)\]
We actually compute $2P$, can then divide by 2.
Fix $t$, we note $P_a$ for $P_{a,t}$.
Notation $P_{-a}$ for $P_{q-a}$ (same since $g$ has order $q$).

Notice that for every $x\in \{g, g^{-1}\}^n$, we have $P_{-a}(X) = P_{a}(X^-1)$.

TODO : needs index i as well.
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\bibliography{biblio.bib}
